---
title: "Big Data in Personal Finance"
author: "Jonathan Ratschat"
date: "3 1 2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning= FALSE, message = FALSE)
```

# Packages

```{r}
library(haven) #working with .dta
library(dplyr) #tidy working
library(GGally) #visualization of correlations
library(MASS) #backward regression
library(survey) #weighted data
library(srvyr) #extension of survey package
library(poliscidata) #model fit statistics for survey package
library(jtools) #standard error calculation for survey package
library(ggplot2) #visualization
library(lmtest) #test if heterogeneity is a problem (Breusch-Pagan test)
library(sandwich) #robustness of standard errors
library(stargazer) #visualization of tables
library(reshape2) #wide to long format, long to wide format
library(tidyr) #tidy messy data
library(HH)
```

# Importing data

```{r}
#Import main survey data
Data <- read_dta("p16i6.dta")
Test <- Data

#Delete J-variables except the J-variables of the independent variables 
#(J-variables indicate missing/imputed data)
Data <- dplyr::select(Data, starts_with("X"), J3913, J3915, Y1, YY1)

#Import core variables (e.g. fin, networth, asset, etc.)
CoreData <- read_dta("rscfp2016.dta")

#Merge main survey data and core variables
Data <- merge(Data, CoreData, by = c("Y1", "YY1"))

#Remove CoreData
rm(CoreData)

#Transform weighting variable to fit single implicate
Data$nwgt <- Data$X42001/(5*20163.523)
summary(Data$nwgt)

#Data using only first implicate
Data <- Data[!duplicated(Data$YY1),]

#Delete households whose stock value was completely missing: 79 observations deleted
#Shadow variables: Codes of 1097 or more indicate that the associated data value 
#was completely missing
Data <- dplyr::filter(Data, J3915 <= 1096)

#Use of survey package for weighting

#Create svydesign for weighting data set
weight <- svydesign(ids = ~1, #no clusters
                    probs = NULL, #no cluster sampling probabilities
                    fpc = NULL, #strata size in the population
                    strata = NULL, #no strata
                    weights = Data$nwgt, data = Data)
```

# Variables

## Stock market particpation

### Stocks

```{r}
#Stocks

#hstocks: Do you (or anyone in your family living here) own any stock which is 
#publicly traded?, yes = 1, no = 0

#X3915 or stocks: total market value of stocks
summary(Data$stocks)
```

### Mutual funds

```{r}
#stmutf: total market value of all stock mutual funds and ETFs

#comutf: total market value of combination mutual funds

#stcomutf: full value if described as stock mutual fund, 1/2 value of combination 
#mutual funds
Data$stcomutf <- Data$stmutf + 0.5 * Data$comutf
summary(Data$stcomutf)

#hstcomutf: does respondent have stock mutual fund or combination mutual fund? yes = 1, no = 0
Data$hstcomutf <- ifelse(Data$stcomutf > 100, 1, 0)
```

### Retirement

```{r}
#Retirement

#reteq: equity in quasi-liquid retirement assets (individual retirement accounts/Keoghs 
#[IRAKH], account-type pensions on current job [THRIFT], future pensions [FUTPEN], 
#currently received accounttype pensions [CURRPEN])
summary(Data$reteq)

#hreteq: does respondent have quasi-liquid retirement assets? yes = 1, no = 0
Data$hreteq <- ifelse(Data$reteq > 100, 1, 0)
```

### Totals

```{r}
#ststcomutf: total market value of stocks and stcomutf
Data$ststcomutf <- Data$stocks + Data$stcomutf

#hststcomutf: has stocks and stcomutf
Data$hststcomutf <- ifelse(Data$ststcomutf > 100, 1, 0)
summary(Data$hststcomutf)

#sttotal: total market value of stocks, stock mutual funds, combination mutual funds and 
#quasi-liquid retirement assets
Data$sttotal <- Data$stocks + Data$stcomutf + Data$reteq
summary(Data$sttotal)

#hsttotal: has stocks, stock mutual funds, combination mutual funds and quasi-liquid 
#retirement assets
Data$hsttotal <- ifelse(Data$sttotal > 100, 1, 0)

summary(Data$hsttotal)
```

## Financial literacy

```{r}
#X7558: Financial literacy diversification, correct answer: 5
#X7559: Financial literacy interest rate, correct answer: 1
#X7560: Financial literacy inflation, correct answer: 5
#finlit: Number of questions right
#baseFinlit: base financial literacy (all three questions right) = 1, not financial 
#literate (two or less questions right) = 0
Data$baseFinlit <- ifelse(Data$finlit == 3, 1, 0)

#as.factor
Data$baseFinlit <- as.integer(Data$baseFinlit)

#X7556 or selfFinlit: Personal finance knowledge, scale 1 to 10, -1 instead of 0
Data$selfFinlit <- ifelse(Data$X7556 == -1, 0, Data$X7556)

#Maybe include:
#X7562: When making saving and investment decisions, some people search for the very 
#best terms while others don't, scale 1 to 10, -1 instead of 0
#X7112 X7113 X7114 X7115 X7116 X7117 X7118 X7119 
#X7120 X7121 X6865 X6866 X6867 X6868 X6869 Please look at this list. 
#What sources of information do you (and your {husband/wife/partner)use to make decisions 
#about saving and investments?(Do you callaround, read newspapers, magazines, material you 
#get in the mail, use information from television, radio, the Internet or advertisements?  
#Do you get advice from a friend, relative, lawyer, accountant, banker, broker, or financial 
#planner?  Or do you do something else?)
```

## Age

```{r}
#X14 or age: Age, numeric
#Create age2
Data$age2 <- Data$age^2

#ageGroups as factor
Data$ageGroups <- ifelse(Data$age >= 18 & Data$age <= 35, 1, 
                         ifelse(Data$age >= 36 & Data$age <= 50, 2, 
                                ifelse(Data$age >= 51 & Data$age <= 65, 3, 4)))
Data$ageGroups <- as.factor(Data$ageGroups)
```

## Sex

```{r}
#X8021 or hhsex: Sex, male = 1, female = 2
#Change to male = 0, female = 1
Data$hhsex <- ifelse(Data$X8021 == 2, 1, 0)
summary(Data$hhsex)

#Create factor variable
Data$hhsex <- as.factor(Data$hhsex)
```

## Marital status

```{r}
#X8023: Marital status, married = 1, living with partner = 2, separated = 3, 
#divorced = 4, widowed = 5, never married = 6, person under 17 = 0
#married: married/living with partner = 1, neither married nor living with partner = 2
#Change to married/living with partner = 1, neither married nor living with partner = 0
Data$married[Data$married == 2] <- 0
summary(Data$married)

#Create factor variable
Data$married <- as.factor(Data$married)
```

# Education

```{r}
#educ or X5913: What is the highest level of school completed or the highest degree you have 
#received? 1st, 2nd, 3rd, or 4th grade = 1, 5th or 6th grade = 2, 7th and 8th grade = 3, 
#9th grade = 4, 10th grade = 5, 11th grade = 6, 12th grade, no diploma = 7, 
#High school graduate - high school diploma or equivalent = 8, Some college but no degree = 9, 
#Associate degree in college - occupation/vocation program = 10, 
#Associate degree in college - academic program = 11, 
#Bachelor's degree (for example: BA, AB, BS) = 12, Master's degree ( for exmaple: MA, MS, 
#MENG, MED, MSW, MBA) = 13, Professional school degree (for example: MD, DDS, DVM, LLB, JD) = 14, 
#Doctorate degree (for example: PHD, EDD) = 15, Less than 1st grade = 0

#Switch -1 with 0
Data$educ[Data$educ == -1] <- 0

#Create educ2
Data$educ2 <- Data$educ^2

#Create factor variable: less than highschool = 1, high school degree and some college = 2, 
#college degree or bachelor degree = 3, master or higher = 4
Data$educFactor <- ifelse(Data$educ >= 0 & Data$educ <= 7, 1, 
                          ifelse(Data$educ == 8 |Data$educ == 9, 2, 
                                 ifelse(Data$educ >= 10 & Data$educ <= 12, 3, 4)))
Data$educFactor <- as.factor(Data$educFactor)
```

## Employed and unemployed

```{r}
#X6670: 1.    *WORKING NOW/SELF-EMPLOYED; job accepted and waiting to start work = 1, 
#TEMPORARILY LAID OFF; seasonal work and not working now = 2, UNEMPLOYED AND LOOKING FOR WORK = 3, 
#STUDENT; job training = 4, *HOMEMAKER; never worked; misc. out of the labor force n.e.c. 
#(including prison) = 5, DISABLED = 6, RETIRED (FULL, PARTIAL, OR TEMPORARY) = 7, 
#ON SICK LEAVE OR MATERNITY LEAVE = 8, VOLUNTEER WORK = 9, ON VACATION/OTHER LEAVE OF ABSENCE = 9, 
#On sabbatical/extended leave and expecting to go back to job = 13, ON STRIKE = 15, OTHER NOT 
#WORKING AND NOT LOOKING FOR WORK = 16

#Create factor variable
Data$X6670 <- as.factor(Data$X6670)

#Create factor indicating employment
Data$employed <- ifelse(Data$X6670 == 1, 1, 0)
summary(Data$employed)

#Create factor indicating unemployment
Data$unemployed <- ifelse(Data$X6670 == 3, 1, 0)
summary(Data$unemployed)

#Create factor indicating being retired
Data$retired <- ifelse(Data$X6670 == 7, 1, 0)
summary(Data$retired)
```

# Sample data set based on financial net worth and income

## Financial net worth

```{r}
#Financial net worth
Data$netfin <- Data$fin - Data$ccbal - Data$install - Data$odebt
summary(Data$netfin)

#Non-financial net worth
Data$netnfin <- Data$nfin - Data$mrthel - Data$resdbt - Data$othloc
summary(Data$netnfin)
```

## Income
```{r}
#income: total income 2015 from all sources, before taxes and other deductions
summary(Data$income)
```

## Sampling

```{r}
Data$netfin <- ifelse(Data$netnfin > 0 & Data$netfin == 0, 1, Data$netfin)

Data <- filter(Data, Data$netfin > 0, Data$income > 0)
```

## Create log of financial net worth and income and control correlation of basic financial literacy and self-assessed financial literacy

```{r}
Data$lognetfin <- log(Data$netfin) 

Data$logincome <- log(Data$income)

#Use of survey package for weighting
#Create svydesign
weight <- svydesign(ids = ~1, #no clusters
                    probs = NULL, #no cluster sampling probabilities
                    fpc = NULL, #strata size in the population
                    strata = NULL, #no strata
                    weights = Data$nwgt, data = Data)

#Calculate correlation
c <- svycor(~baseFinlit + selfFinlit, design = weight, sig.stats = TRUE)
c$cors
c$p.values
c$std.err
```

## Real estate

```{r}
#houses: value of primary residence
summary(Data$houses)

#oresre: value of other residential real estate (e.g. vacation homes)
summary(Data$oresre)

#housing value: value of primary residence + other residential real estate
Data$housing <- Data$houses + Data$oresre
Data$housing[Data$housing == -1] <- 0
summary(Data$housing)

#nnresre: net equity in nonresidential real estate
summary(Data$nnresre)

#thousing: total housing value
Data$thousing <- Data$housing + Data$nnresre
summary(Data$thousing)

#thousingRatio: total housing value relative to asset value
Data$thousingRatio <- Data$thousing / Data$asset
Data$thousingRatio <- ifelse(is.na(Data$thousingRatio), 0, Data$thousingRatio)
summary(Data$thousingRatio)

#Binary variable indicating home ownership
Data$hthousing <- ifelse(Data$thousing > 10000, 1, 0)
Data$hthousing <- as.factor(Data$hthousing)
```

## Business

```{r}
#bus: business interests; for businesses where the HH has an active interest, value is net equity if 
#business were sold today, plus loans from HH to business, minus loans from business to HH not 
#previously reported, plus value of personal assets used as collateral for business loans that 
#were reported earlier; for businesses where the HH does not have an active interest, market value of the interest
summary(Data$bus)

#busRatio: business value relative to asset value
Data$busRatio <- Data$bus / Data$asset
Data$busRatio <- ifelse(is.na(Data$busRatio), 0, Data$busRatio)
summary(Data$busRatio)

#Binary variable indicating home ownership
Data$hbus <- ifelse(Data$bus > 10000, 1, 0)
Data$hbus <- as.factor(Data$hbus)

Test <- filter(Data, bus < 10000 & bus > 0)
summary(Test$netfin)
```

## Risk attitude

```{r}
#X3014:  Which of the statements on this page comes closest to the amount of financial risk that 
#you (and your {husband/wife/partner}) are willing to take when you save or make investments? 
#Take substantial financial risks expecting to earn substantial returns = 1, 
#Take above average financial risks expecting to earn above average returns = 2, 
#Take average financial risks expecting to earn average returns = 3, 
#Not willing to take any financial risks = 4
Data$finRisk <- Data$X3014
Data$finRisk <- as.factor(Data$finRisk)

summary(Data$finRisk)

Data$finRiskNumeric <- ifelse(Data$X3014 == 1, 4, 
                              ifelse(Data$X3014 == 2, 3, 
                                     ifelse(Data$X3014 == 3, 2, 1)))

prop.table(table(Data$finRisk))
prop.table(table(Data$finRiskNumeric))
```

# Descriptives

## Table 1 - Sample statistics

```{r}
#Describtive table: mean, SD

weightsrvyr <- Data %>%
   as_survey_design(weights = nwgt)

#Create svydesign
weight <- svydesign(ids = ~1, #no clusters
                    probs = NULL, #no cluster sampling probabilities
                    fpc = NULL, #strata size in the population
                    strata = NULL, #no strata
                    weights = Data$nwgt, data = Data)

rowAge <- weightsrvyr %>% 
  srvyr::summarise(X14 = survey_mean(X14))
rowAge <- cbind(round(rowAge[,1], digit = 0), rowAge[,2])

colnames(rowAge) <- c("Mean", "Std. Err.")

rowAge1 <- svymean(~ageGroups, design = weight)
rowAge1 <- as.data.frame(rowAge1)
rowAge1

colnames(rowAge1) <- c("Mean", "Std. Err.")

rowMen <- svymean(~hhsex, design = weight)
rowMen <- as.data.frame(rowMen)
rowMen <- rowMen[1,]

colnames(rowMen) <- c("Mean", "Std. Err.")

rowMarried <- svymean(~married, design = weight)
rowMarried <- as.data.frame(rowMarried)
rowMarried <- rowMarried[2,]

colnames(rowMarried) <- c("Mean", "Std. Err.")

rowEduc <- svymean(~educFactor, design = weight)
rowEduc <- as.data.frame(rowEduc)

colnames(rowEduc) <- c("Mean", "Std. Err.")

rowEmployed <- svymean(~employed, design = weight)
rowEmployed <- as.data.frame(rowEmployed)
rowEmployed <- rowEmployed[1,]

colnames(rowEmployed) <- c("Mean", "Std. Err.")

rowUnemployed <- svymean(~unemployed, design = weight)
rowUnemployed <- as.data.frame(rowUnemployed)
rowUnemployed <- rowUnemployed[1,]

colnames(rowUnemployed) <- c("Mean", "Std. Err.")

rowRetired <- svymean(~retired, design = weight)
rowRetired <- as.data.frame(rowRetired)

colnames(rowRetired) <- c("Mean", "Std. Err.")

rowIncome <- weightsrvyr %>% 
  srvyr::summarise(income = survey_mean(income))
rowIncome <- cbind(round(rowIncome[,1], digit = 0), rowIncome[,2])

colnames(rowIncome) <- c("Mean", "Std. Err.")

rowNetfin <- weightsrvyr %>% 
  srvyr::summarise(netfin = survey_mean(netfin))
rowNetfin <- cbind(round(rowNetfin[,1], digit = 0), rowNetfin[,2])

colnames(rowNetfin) <- c("Mean", "Std. Err.")

Part1 <- rbind(rowAge, rowAge1, rowMen, rowMarried, rowEduc, rowEmployed, 
               rowUnemployed, rowRetired, rowIncome, rowNetfin)

Part1 <- as.data.frame(Part1)

Part1

rownames(Part1) <- c("Age",
                     "Age: 18 to 35", 
                     "Age: 36 to 50", 
                     "Age: 51 to 65", 
                     "Age: over 65",  
                     "Men", 
                     "Married or living with a partner", 
                     "Education: lower than high school", 
                     "Education: high school graduate or some college but no degree", 
                     "Education: college degree or bachelor's degree", 
                     "Education: master's degree or higher", 
                     "Employed", 
                     "Unemployed", 
                     "Retired", 
                     "Total income 2015 (USD)", 
                     "Net financial wealth - end of 2015 (USD)")

Part1
```

```{r}
#Describtive table: min, max

weightsrvyr <- Data %>%
   as_survey_design(weights = nwgt)

rowAge <- weightsrvyr %>% 
  srvyr::summarise(X14 = min(X14), X14 = max(X14))

rowAge1 <- data.frame("Min" = 1, "Max" = 4)
rowAge1

rowAge2 <- data.frame("Min" = 1, "Max" = 4)
rowAge2

rowAge3 <- data.frame("Min" = 1, "Max" = 4)
rowAge3

rowAge4 <- data.frame("Min" = 1, "Max" = 4)
rowAge4

colnames(rowAge) <- c("Min", "Max")

rowMen <- data.frame("Min" = 0, "Max" = 1)
rowMen

rowMarried <- data.frame("Min" = 0, "Max" = 1)
rowMarried

rowEduc1 <- data.frame("Min" = 1, "Max" = 4)
rowEduc1

rowEduc2 <- data.frame("Min" = 1, "Max" = 4)
rowEduc2

rowEduc3 <- data.frame("Min" = 1, "Max" = 4)
rowEduc3

rowEduc4 <- data.frame("Min" = 1, "Max" = 4)
rowEduc4

rowEmployed <- data.frame("Min" = 0, "Max" = 1)
rowEmployed

rowUnemployed <- data.frame("Min" = 0, "Max" = 1)
rowUnemployed

rowRetired <- data.frame("Min" = 0, "Max" = 1)
rowRetired

rowIncome <- weightsrvyr %>% 
  srvyr::summarise(income = min(income), income = max(income))
rowIncome <- cbind(rowIncome[,1], round(rowIncome[,2], digit = 0))

colnames(rowIncome) <- c("Min", "Max")

rowNetfin <- weightsrvyr %>% 
  srvyr::summarise(netfin = min(netfin), netfin = max(netfin))

colnames(rowNetfin) <- c("Min", "Max")

Part2 <- rbind(rowAge, rowAge1, rowAge2, rowAge3, rowAge4, rowMen, 
               rowMarried, rowEduc1, rowEduc2, rowEduc3, rowEduc4, 
               rowEmployed, rowUnemployed, rowRetired, rowIncome, rowNetfin)

Part2 <- as.data.frame(Part2)

rownames(Part2) <- c("Age","Age: 18 to 35", 
                     "Age: 36 to 50", 
                     "Age: 51 to 65", 
                     "Age: over 65",  
                     "Men", 
                     "Married or living with a partner", 
                     "Education: lower than high school", 
                     "Education: high school graduate or some college but no degree", 
                     "Education: college degree or bachelor's degree", 
                     "Education: master's degree or higher", 
                     "Employed", 
                     "Unemployed", 
                     "Retired", 
                     "Total income 2015 (USD)", 
                     "Net financial wealth - end of 2015 (USD)")

Part2

summary(Data$netfin)
```

```{r}
#Create table 1
Table1 <- cbind(Part1, Part2)
Table1

#Export it as word document
stargazer(Table1, 
          title="Descriptive Statistics",
          summary = FALSE,
          type = "html", 
          out= "star_descriptive_1.doc")

#Text for Table 1
#Table 1: Sample statistics
#This table contains sample statistics for 6,169 respondents in SCF 2016.
#Source: SCF 2016, data are weighted
```

## Table 2

```{r}
#Define weight design
weightsrvyr <- Data %>%
   as_survey_design(weights = nwgt)

#Create table for Table 2
Table2 <- svyby(formula = ~baseFinlit, 
             by = ~hhsex,
             design = weightsrvyr, 
             FUN = svymean, 
             na.rm = TRUE, 
             keep.names = FALSE)

#Make sure that Table2 is a data frame
(Table2 <- as.data.frame(Table2))

#Change variable output names
Table2$hhsex <- factor(Table2$hhsex, levels = c("0", "1"),
                       labels = c("Men", "Women"))

Table2

#Round results
Table2$baseFinlit0 <-  1 - Table2$baseFinlit
Table2$baseFinlit1 <- Table2$baseFinlit

#Select columns
Table2 <- Table2 %>% select(hhsex, baseFinlit0, baseFinlit1)

Table2

#Change column names
colnames(Table2) <- c("Gender", "Illiterate", "Literate")

#Create row for total financial literacy
rowTotal <- svymean(~baseFinlit, design = weightsrvyr)
rowTotal <- as.data.frame(rowTotal)
rowVariable <- c("Illiterate", "Literate")
rowTotal <- cbind(rowTotal, rowVariable)
rowTotal <- spread(rowTotal, rowVariable, mean)
rowTotal <- rowTotal %>% select("Illiterate", "Literate")
rowDescription <- "Total"
rowTotal <- cbind(rowDescription, rowTotal)
colnames(rowTotal) <- c("Gender", "Illiterate", "Literate")
Table2 <- rbind(Table2, rowTotal)

percent <- function(x, digits = 2, format = "f", ...) {
  paste0(formatC(100 * x, format = format, digits = digits, ...), "%")
}

Table2$Illiterate <- percent(Table2$Illiterate)
Table2$Literate <- percent(Table2$Literate)

Table2

#Export it as word document
stargazer(Table2, 
          title="Descriptive Statistics",
          summary = FALSE,
          type = "html", 
          out= "star_descriptive_2.doc")
```

## Info 1 - Stock market particpation based on financial literacy

```{r}
#Define weight design
weightsrvyr <- Data %>%
   as_survey_design(weights = nwgt)

#Create table for Graph 1
Info1 <- svyby(formula = ~hsttotal, 
             by = ~baseFinlit,
             design = weightsrvyr, 
             FUN = svymean, 
             na.rm = TRUE, 
             keep.names = FALSE)

Info1 <- as.data.frame(Info1)
Info1$hsttotal <- round(Info1$hsttotal, digits = 4)

Info1

svyttest(hsttotal~baseFinlit, design = weightsrvyr)
```

## Graph 1 - Types of financial assets and stock market participation over financial literacy

```{r}
#Define weight design
weightsrvyr <- Data %>%
   as_survey_design(weights = nwgt)

#Create table for Graph 5
Graph1 <- svyby(formula = ~ hststcomutf + hreteq + hsttotal, 
             by = ~baseFinlit,
             design = weightsrvyr, 
             FUN = svymean, 
             na.rm = TRUE, 
             keep.names = FALSE)

#Make sure that Graph1 is a data frame
Graph1 <- as.data.frame(Graph1)
Graph1

#Transform Graph1 to long format
Graph1Long <- melt(Graph1, id = "baseFinlit", 
                   measure = c("hststcomutf", "hreteq", "hsttotal"))
Graph1Long$variable <- factor(Graph1Long$variable, 
                              levels = c("hststcomutf", "hreteq", "hsttotal"), 
                  labels = c("Stocks and stock mutual funds", "Quasi-liquid retirement assets", "Total"))
Graph1Long

Graph1Long$baseFinlit <- c("Illiterate", "Literate", "Illiterate", "Literate", "Illiterate", "Literate")

#Create Graph 1 Long with ggplot2
ggplot(data = Graph1Long, aes(x = baseFinlit,y = value, fill = variable)) + 
  geom_col(fill = c("lightgrey","grey", "lightgrey","grey", "lightgrey","grey")) +
  facet_wrap(~variable) +
  xlab("Financial literacy") + 
  ylab("Stock market participation") +
  scale_y_continuous(breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), 
                     labels = c("0%", "10%", "20%", "30%", "40%", "50%", "60%", "70%", "80%", "90%", "100%"), 
                     expand = c(0,0), limits=c(0,1)) +
  expand_limits(y = 0) +
    theme(## plotregion
        panel.background = element_rect(fill = "white",
                                colour = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "lightgrey"),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(0.25, "lines"),
        ## axis line
        axis.line = element_line(colour = "black", 
                      size = 0.5, linetype = "solid"),
        ##background of fill
        strip.background = element_rect(fill="#f2f2f2"),
        ##delete legend
        legend.position = "none"
  )

#Save Graph1
Graph1
ggsave("Graph1.png", width = 20, height = 10, units = "cm")

#t-Test
svyttest(hststcomutf~baseFinlit, design = weightsrvyr)
```

## Graph 2 - Financial literacy and stock market participation over levels of educa-tion

```{r}
#Define weight design
weightsrvyr <- Data %>%
   as_survey_design(weights = nwgt)

#Create table for Graph 2
Graph2 <- svyby(formula = ~hsttotal, 
             by = ~educFactor + baseFinlit,
             design = weightsrvyr, 
             FUN = svymean, 
             na.rm = TRUE, 
             keep.names = FALSE)

#Make sure that Graph1 is a data frame
Graph2 <- as.data.frame(Graph2)

#Change variable format so that graph works
Graph2$educFactor <- as.numeric(Graph2$educFactor)
Graph2

Graph2$baseFinlit <- c("Illiterate", "Illiterate", "Illiterate", 
                       "Illiterate", "Literate", "Literate", "Literate", "Literate")

#Create Graph 2 with ggplot2
ggplot(data = Graph2, aes(x = educFactor,y = hsttotal)) + 
  geom_line(aes(linetype = baseFinlit)) +
  geom_point() +
  xlab("Education level") + 
  ylab("Stock market participation") +
  labs(linetype = "Financial literacy") +
  scale_y_continuous(breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), 
                     labels = c("0%", "10%", "20%", "30%", "40%", "50%", "60%", "70%", "80%", "90%", "100%"), 
                     expand = c(0,0), limits=c(0,1)) +
  expand_limits(y = 0) +
  theme(## legend
    legend.title.align = 0.5,
        legend.position = "bottom",
        legend.direction = NULL,
        legend.justification = "center",
        legend.box = "vertical",
        ## plotregion
        panel.background = element_rect(fill = "white",
                                colour = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "lightgrey"),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(0.25, "lines"),
        ## axis line
        axis.line = element_line(colour = "black", 
                      size = 0.5, linetype = "solid")
  )

#Save Graph2
ggsave("Graph2.png", width = 20, height = 10, units = "cm")
```

## Graph 3 - Financial literacy and stock market participation over percentiles of total net worth

```{r}
#Define weight design
weightsrvyr <- Data %>%
   as_survey_design(weights = nwgt)

#Create table for Graph 3
Graph3 <- svyby(formula = ~hsttotal, 
             by = ~baseFinlit + nwcat,
             design = weightsrvyr, 
             FUN = svymean, 
             na.rm = TRUE, 
             keep.names = FALSE)

#Make sure that Graph1 is a data frame
Graph3 <- as.data.frame(Graph3)

#Calculate percentage change column
Graph3 %>%
  mutate(pct_change = (hsttotal/lag(hsttotal) - 1) * 100)

Graph3$baseFinlit <- c("Illiterate", "Literate", "Illiterate", "Literate", 
                       "Illiterate", "Literate", "Illiterate", "Literate", 
                       "Illiterate", "Literate")

#Create Graph 3 with ggplot2
ggplot(data = Graph3, aes(x = nwcat,y = hsttotal)) + 
  geom_line(aes(linetype = baseFinlit)) +
  geom_point() +
  xlab("Net worth percentile groups") + 
  ylab("Stock market participation") +
  labs(linetype = "Financial literacy") +
  scale_y_continuous(breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), 
                     labels = c("0%", "10%", "20%", "30%", "40%", "50%", "60%", "70%", "80%", "90%", "100%"), 
                     expand = c(0,0), limits=c(0,1)) +
  expand_limits(y = 0) +
  theme(## legend
    legend.title.align = 0.5,
        legend.position = "bottom",
        legend.direction = NULL,
        legend.justification = "center",
        legend.box = "vertical",
        ## plotregion
        panel.background = element_rect(fill = "white",
                                colour = "white"),
        panel.border = element_blank(),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "lightgrey"),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(0.25, "lines"),
        ## axis line
        axis.line = element_line(colour = "black", 
                      size = 0.5, linetype = "solid")
  )

#Save Graph3
ggsave("Graph3.png", width = 20, height = 10, units = "cm")
```
 
# Probit regression

```{r}
#"Normal" probit regression with weight variable
probit1 <- glm(hsttotal ~ baseFinlit + selfFinlit + ageGroups + hhsex + married + 
                 educFactor + logincome + lognetfin + hthousing + hbus + finRisk, 
               data = Data, weight = nwgt, family = binomial(link = "probit"))
summary(probit1)
```

```{r}
#Probit with weight variable with robust standard errors using sandwish package
probit1weight1 <- coeftest(probit1, vcov = vcovHC(probit1))
summary(probit1weight1)
```

```{r}
#Probit with weight variable with robust standard errors using survey package

weight <- svydesign(ids = ~1, #no clusters
                    probs = NULL, #no cluster sampling probabilities
                    fpc = NULL, #strata size in the population
                    strata = NULL, #no strata
                    weights = Data$nwgt, data = Data)

##Probit without financial literacy measure
probit1weight2 <- svyglm(hsttotal ~ ageGroups + hhsex + married + educFactor + 
                           logincome + lognetfin + hthousing + hbus + finRisk, 
                         design = weight, family = binomial(link = "probit"))
summary(probit1weight2)

##Probit with financial literacy measure
probit1weight3 <- svyglm(hsttotal ~ baseFinlit + selfFinlit + ageGroups + hhsex + married + 
                           educFactor + logincome + lognetfin + hthousing + hbus + finRisk, 
                         design = weight, family = binomial(link = "probit"))
summary(probit1weight3)

##Probit with financial literacy measure
probit1weight4 <- svyglm(hsttotal ~ baseFinlit + selfFinlit + ageGroups + hhsex + married + 
                           educFactor + logincome + lognetfin + hthousing + hbus + finRiskNumeric, 
                         design = weight, family = binomial(link = "probit"))
summary(probit1weight3)

#Get Pseudo R-Squared
psrsq(probit1weight2, method = c("Cox-Snell"))

psrsq(probit1weight3, method = c("Cox-Snell"))
```

```{r}
#Plot probit1
stargazer(probit1weight3, probit1weight4, 
          single.row = TRUE,
          covariate.labels = c("Basic financial literacy", 
                               "Self-assessed financial literacy", 
                               "Age: 36 to 50", "Age: 51 to 65", 
                               "Age: over 65", 
                               "Men",
                               "Married or living with partner", 
                               "Education: high school graduate or some college but no degree", 
                               "Education: college degree or bachelor's degree", 
                               "Education: master's degree or higher", 
                               "Ln(Total Income)", "Ln(Net financial wealth)", 
                               "Real estate ownership", "Business ownership", 
                               "Risk attitude: take above average financial risks", 
                               "Risk attitude: take average financial risks", 
                               "Risk attitude: not willing to take any financial risks"),
          type="html",
          out="star_probit1.doc")
```

### SAVE FOR LATER ### NOT NEEDED ANYMORE

# Multiple linear regression

```{r}
#"Normal" multiple linear regression
mod1 <- lm(sttotal ~ baseFinlit + ageGroups + hhsex + married + educFactor 
           + income + netfin + hthousing + hbus + finRisk, 
           data = Data, weights = nwgt)
summary(mod1)

#Step-wise backward regression
mod2 <- stepAIC(mod1, direction = "both", trace = FALSE)
summary(mod2)

#Multiple linear regression with weight variable with robust standard errors using sandwish package
modweight1 <- coeftest(mod1, vcov = vcovHC(mod1))
summary(modweight1)
```

```{r}
#Multiple linear regression with robust standard errors using survey package (weight in design included)
weight <- svydesign(ids = ~1, #no clusters
                    probs = NULL, #no cluster sampling probabilities
                    fpc = NULL, #strata size in the population
                    strata = NULL, #no strata
                    weights = Data$nwgt, data = Data)

modweight2 <- svyglm(sttotal ~ageGroups + hhsex + married + educFactor + 
                       logincome + lognetfin + hthousing + hbus + finRisk, 
                     design = weight)
summary(modweight2)

modweight3 <- svyglm(sttotal ~ baseFinlit + ageGroups + hhsex + married + 
                       educFactor + logincome + lognetfin + hthousing + hbus + finRisk, 
                     design = weight)
summary(modweight3)

#Get R-Squared and adjusted R-Squared
fit.svyglm(modweight3, digit = 4)
```

```{r}
#Plot
stargazer(modweight2, modweight3, 
          type="html",
          covariate.labels = c("Financial literacy", 
                               "Age: 36 to 50", 
                               "Age: 51 to 65", 
                               "Age: over 59", 
                               "Men", 
                               "Married or living with partner", 
                               "Education: high school graduate or some college but no degree", 
                               "Education: college degree or bachelor's degree", 
                               "Education: master's degree or higher", 
                               "Total Income 2015", 
                               "Net wealth - end of 2015", 
                               "Real estate ownership", 
                               "Business ownership", 
                               "Risk attitude"),
          out="star_linear.doc")
```

# Test for Multicollinearity

Multicollinearity is when there is correlation between predictors (i.e. independent variables) 
in a model. Its presence can adversely affect our regression results.

```{r}
#Create correlation matrix
#ggcorr(dplyr::select(Data, hsttotal,baseFinlit, ageGroups, hhsex, 
#married, educFactor1, logincome, lognetfin, hthousing, hbus, finRisk), label = TRUE)
```

In the next step, we test the variance inflation factor (VIF). The VIF estimates how much the 
variance of a regression coefficient is inflated due to multicollinearity in the model. It 
quantifies the severity of multicollinearity in an OLS regression analysis.

```{r}
#Calculate VIF per independent variable in model
#vif(mod2)

#Calculate mean of VIFs
#mean(vif(mod2))
```

An average mean between 1 and 5 indicates a moderate correlated result. In general, a VIF 
above 10 indicates high correlation and is cause for concern. Some authors suggest a more 
conservative level of 2.5 or above. Since our average VIF is under 2.5, we assume that 
multicollinearity is not considered as a problem in step.model.

# Test for Homoscedasticity

Homoscedasticity means that the residuals do not have a constant variance. We need to 
check if residuals are normally distributed.

```{r}
#shapiro.test(rstandard(mod1))
```

Shapiro-Wilk-Test of standardized residuals is significant. Residuals differ significantly 
from normal distribution; assumption of homoscedasticity is violated.

The absence of a normal distribution signifies only that the F- and t-tests are not meaningfully 
applicable. The estimated regression coefficients are still unbiased.

# Expected value of zero for the residuals E(e~i~) = 0

```{r}
#wilcox.test(rstandard(mod1))
```

Zero hypothesis cannot be rejected.

The standardized residual is the residual divided by its standard deviation. We use it to show 
the difference of mean of residuals from 0 graphically.

```{r}
#boxplot(rstandard(mod1)) 
#abline(h=0)
```





